\LARGE \bf How Children Perceive and Interact with a Robot that Behaves Unexpectedly - The Domino Experiment

### 2014-09-08 paper.tex
Work on data collection + findings

\usepackage{subfig}
\usepackage{hyperref}
\usepackage{graphicx}
\graphicspath{{figures/}}
\usepackage{multirow}

@@

\subsection{Data collection}
\subsubsection{Action coding}

@@

In total, we obtained 2354 distinct actions which summed up to a total
annotation duration of 145 minutes. On average, one child accounted for 92
actions (SD=23).

@@

\subsubsection{Semi-structured interviews}

\paragraph{Anthropomorphism index}

@@

\subsubsection{Manipulation check: Negligable impact of the conditions}

>>>>

\usepackage{subfig}
\usepackage{hyperref}
\usepackage{graphicx}
\graphicspath{{figures/}}

\usepackage{booktabs}

@@

\subsection{Data collection}

We have obtained and analysed two types of data. The \textbf{children's
behaviours toward the robot} (\ie the child-robot interactions) have been
captured in the video recordings by annotating a set of actions, described
below. The children's \textbf{perception of the
robot} has been captured through audio-recorded semi-structured interviews.

\subsubsection{Interaction: Action coding}

@@

In total, we obtained 2354 distinct actions which summed up to a total
annotation duration of 145 minutes. On average, one child accounted for 92
actions (SD=23). Figure~\ref{fig:domino-time-all} shows the distribution of
actions over the different runs, in any condition.

\begin{figure*}[ht!] 
    \centering 
    \includegraphics[width=0.8\linewidth]{domino-time-all.pdf} 
    \caption{\small \textbf{Number and type of actions for each run} (average for
    one child per minute). Generally, the number of actions does not decrease over
    time (from run 1.1 to run 2.9). The first 7 runs correspond to the
    \textit{expected phase}, the second 7 runs correspond to the \textit{unexpected
    phase}. Especially during run 2.3, the first time when the robot showed an
    unexpected behavior, children tended to \textit{look} more at the experimenter.
    During the unexpected phase, also \textit{talk} and \textit{gesture} seem to be
    increased.}

    \label{fig:domino-time-all} 
\end{figure*}

@@

\subsubsection{Perception: semi-structured interviews}

It is not easy to interview young children: they have a short attention span, we
(the experimenters) are strangers to them, and they are in an unfamiliar
environment. The pre-interview served to ``break the ice'' and during the first
playful interaction phase, most children familiarized themselves with the
situation.

Also, children do not articulate themselves and understand questions like
adults. We tested our interview structure and the script in a pilot-study with
three young children, and were assisted in formulating our questions by a
pedagogue and father of four. This helped us adapting our language and
question-style to the 4-5 year old participants. We set up the interviews like a
casual conversation / discussion, so we did not separate the two children in
order to keep the situation natural. We used casual and informal language
adapted to the age of the children, and encouraged them to justify their answers
and tell us more details by asking, for example, \textit{How do you know ... ?}
or \textit{Tell me, what have you observed when ... ?}. We paid attention to not
``put words in children's mouth''. Consequently, though we re-phrased and
repeated some questions, we never forced children to give an answer, and
accepted when they said they would not know or when they did not respond at all.

\begin{table}[h]
\centering
\footnotesize
\begin{tabular}{p{0.9\linewidth}}
    \toprule
    \textbf{Construct \emph{expectation}} \\
    \midrule

    \emph{How do you imagine a robot?} \\
    \emph{What could it look like?} \\
    \emph{Have you ever seen a robot before?} \\

    %%
    \toprule
    \textbf{Construct \emph{impression}} \\
    \midrule

    \emph{When you first saw R, what did you think?} \\
    \emph{Is R a robot? How do you know?} \\
    \emph{Did you expect R would come over to you when you call it?} \\
    \emph{What happened when you put the domino tile in the box?} \\

    %%
    \toprule
    \textbf{Ascribe intention} \\
    \midrule

    \emph{Do you think R could go out the door all by itself?} \\	
    \emph{Does R always obey / come over to you?} \\
    \emph{Could R do something silly?} \\
    \emph{Why did R not come over to you when you called it?} \\

    %%
    \toprule
    \textbf{Ascribe cognitive abilities} \\
    \midrule

    \emph{Here is a domino tile. Do you think R can see it?} \\ 
    \emph{When I say \textit{``Hello R!''}, do you think R can hear it?} \\

    %%
    \toprule
    \textbf{Ascribe emotional state} \\
    \midrule

    \emph{Does R have feelings? Can R be happy or sad sometimes?} \\

    %%
    \toprule
    \textbf{Social acceptance} \\
    \midrule

    \emph{Do you like R? Why (not)?} \\
    \emph{What do you (not) like about it?} \\
    \emph{Would you like to have R at home?} \\

    %%
    \toprule
    \textbf{Companionship} \\
    \midrule

    \emph{Could R be your friend? Why (not)?}\\

    %%
    \toprule
    \textbf{Ascribe moral standing} \\
    \midrule

    \emph{Assume you go on a holiday for two weeks. Is it alright to leave R
    alone at home? Why (not)?} \\

    \bottomrule

\end{tabular}

    \caption{\small Questions used during the semi-structured interviews
    with children. The Ranger robot toy box is abbreviated with ``R''.}

    \label{tab:domino-questions}

\end{table}

We generally needed to keep the interviews as short as possible. In designing
our interview script and selecting relevant questions, we took inspiration from
previous work on child-robot interaction and children's perception of robots
\cite{kahn_jr._robotic_2006,weiss_i_2009,leite_influence_2013}. For instance, we
applied and adapted some of the ``constructs'' and example questions from the
questionnaires used in \cite{kahn_jr._robotic_2006} and \cite{weiss_i_2009}. The
authors evaluated children's perception of the robotic dog AIBO after the
children had played with the robot. A ``construct'' addresses a specific factor
(topic) that can be measured by several questions. For instance, the construct
``cognitive abilities'' (called ``cognition'' in \cite{weiss_i_2009}) considers
the robot's ability to hear and to see (perceptual skills), as attributed by the
children. The construct ``moral standing'' and the respective question was taken
from \cite{kahn_jr._robotic_2006}.\footnote{According to
\cite{kahn_jr._robotic_2006}, \textit{moral} refers to considerations based on
an artifact's physical or psychological welfare, and virtue (whether the
artifact deserves care). An attribution of moral standing reflects, for
instance, that the robot engenders moral regard, is morally responsible,
blameworthy, has rights or deserves respect.}			Similarly, we grouped
questions according to specific constructs that they evaluate (see
Table~\ref{tab:domino-questions}). This was an adaption and extension of the
questions and constructs used in previous work.\\

With several recurring questions in the first and second interview, we wanted to
see the differences in children's perception of the correctly behaving and
unexpectedly behaving robot. We planned to use these two interviews as a
within-subject measurement, however, this did not work out very well because
children's responses were not always accurate, not comparable one by one, and
children did not always give an answer.

%	Even when children reply that they do not know, this is interesting. Then we either do not ask a good question or they are really uncertain about it.

Similar to how we processed interviews in the previous studies, we qualitatively
transcribed interviews, \ie we did not craft a full word-by-word transcript but
noted down any statement that was useful and relevant. This was organized in a
spreadsheet, so to obtain an overview of the different replies to a question.
The interviews were analyzed in a qualitative manner.

\subsubsection{Engagement and anthropomorphism index}

we can investigate how far these two types of data can
be used to understand children's engagement in the interaction and how far they
anthropomorphize the robot. We would like to develop a toolkit that considers
both children's perception and interaction, in order to measure anthropomorphism
(as a special type of human-like engagement with the robot).

There are different possibilities to measure engagement in HRI, depending on the
specific research question and context. Metrics to measure behavioral engagement
in the interaction include, for instance, conversation analysis (\eg used in
\cite{short_no_2010}) or general attention analysis. These can be studied by
analyzing interaction videos in terms of head movement, eye tracking and gesture
/ body movement analysis (\eg in \cite{sidner_explorations_2005}).
Post-measurements to measure emotional engagement can be questionnaires that try
to assess constructs like the perceived presence and involvement in the
interaction. An example is the \textit{Interactive Experience Questionnaire}
(originally developed by \cite{lombard_measuring_2000}) of which adapted
versions were used in
\cite{kidd_effect_2004,bainbridge_effect_2008,short_no_2010}.

A mix of several methods has been used in a long-term interaction study with 8-9
year old children, \cite{leite_long-term_2013}. The author measured engagement
through video observations (by analyzing the amount of time that children spent
looking at the robot), interviews, and questionnaires. The interviews were
semi-structured, containing initial yes-or-no questions followed by open-ended
questions that allowed children to justify and elaborate their answers. We have
chosen a similar approach that considers both children's behavioral and
emotional engagement.

With 4-5 year old children, however, we cannot use rating scales to ask them
about constructs as abstract as social presence or how much they felt involved
in the interaction. Therefore we have attempted to combine \emph{interaction}
data with the results of the semi-structured interviews.  Several of the
aforementioned coded actions can reflect engagement: \textit{explore, misuse,
gesture, touch, show,} and \textit{talk}. Also, \textit{look} at the
experimenter can be considered as engagement in the interaction: We assume that
children look at the experimenter when they are surprised and seek for help.
This behavior reflects that they want to make sense of the robot's behavior and
that they notice it as ``strange''. Further, we can take into account how they
refer to Ranger and describe the robot (as well as their experience) in the
interviews.  Specifically, we used the constructs \textit{ascribe intention,
    ascribe cognitive abilities, ascribe emotional state}, and \textit{ascribe
    moral standing} (Table~\ref{tab:domino-questions}) to assess how far they
    anthropomorphize the robot.

To account for the fact that anthropomorphism arises in an interaction, we try
to bring both children's perception of the robot (post-measurement) and their
behavior toward it (in-the-moment measurement) together. By doing so, we would
like to obtain a qualitative \textbf{anthropomorphism
index}~\cite{fink2014dynamics}. We quantify the index by giving 1 point for each
anthropomorphic perception of the robot (max.  13 points) and for specific kinds
of human-like behavior toward the robot (max.  3 points). The index included the
following aspects:\\

\textbf{Perception: \textit{(max. 13 points)}}
\begin{itemize}
	\item Ascribe \textbf{mental states / feelings} to Ranger: \textit{0-4 points}\\
	(2 points for agreeing that Ranger can be happy or sad; 2 points for attributing Ranger with hunger or tiredness)
	\item Ascribe \textbf{cognitive abilities / intention} to Ranger: \textit{0-4 points}\\ 
	(each 0.5 points for ascribing seeing and hearing ability; 1 point for agreeing that Ranger can go out the door by itself; 1 point for disagreeing that Ranger always obeys; 1 point for agreeing that Ranger can do something silly)
	\item Ascribe \textbf{sociality / companionship} to Ranger: \textit{1 point}\\ 
	(1 point for agreeing that Ranger can be a friend)
	\item Ascribe \textbf{moral standing} to Ranger: \textit{1 point}\\ (1 point for disagreeing that Ranger be left alone at home)
	\item Other \textbf{anthropomorphic statements}: \textit{0-3 points}\\ 
	(1 point for anthropomorphic reason for Ranger's misbehavior; 2 points for anthropomorphic reason for not leaving Ranger alone) 
\end{itemize}

\textbf{Behavior: \textit{(max. 3 points)}}
\begin{itemize}
	\item Use of \textbf{direct speech} toward Ranger: \textit{1 point}\\
	(\textit{calling} the robot to come over is not considered)
	\item Use of \textbf{polite formulations} toward Ranger: \textit{1 point}\\ 
	(\eg saying \textit{``thank you Ranger''} or \textit{``please Ranger ...''} or \textit{``goodbye''})
	\item Use of \textbf{social or pointing gestures} toward Ranger: \textit{1 point}\\ 
	(\eg waving at the robot, nodding)
\end{itemize}

There are certainly limitations to this scoring scheme. For instance, the
balance between perception and behavior aspects is questionable (13 to 3
points). Also, we did not consistently assign 1 point to each item, but assigned
points between 0.5 and 2 points. We did this because we found that different
items reflect a higher level of anthropomorphic perception of the robot than
others (for instance, ascribing the ability to see an hear was suggested by our
study setup, and we cannot be sure that it really reflects anthropomorphism).
There are possibilities of refinement of this scoring scheme. For instance, a
more systematic weighting procedure and the identification of all relevant
factors would be an interesting future achievement. We could introduce a formula
to calculate the anthropomorphism index \textit{$\alpha$} by taking the sum of
the various items ($x_1, x_2, x_3...$) and different weights ($a, b, c ...$) to
balance them: $\alpha$ \textit{=} $a x_1$ \textit{+} $b x_2$ \textit{+} $c x_3
...$~. Machine learning techniques could be used to optimize such a calculation
of the anthropomorphism index.

@@

\paragraph{Moral standing}

Inspired from the questionnaire used by \cite{kahn_jr._robotic_2006}, we asked
children if it would be alright to leave Ranger alone at home (\eg during two
weeks when they go on a vacation) (Figure~\ref{fig:domino-leave-alone}). The great
majority of 20 children responded negatively. Asked why, children gave a variety
of answers that we classified into 7 categories. With 5 replies, the most common answer was
that the robot ``could do something silly'' (\emph{``faire une bêtise''}). Some
other children simply answered they would like to take it with them. Others
were afraid that the Ranger ``would not find its way'' or ``may be taken
away by someone''. 2 children believed Ranger is sad when left alone and 1
child responded the robot would try to escape. Interestingly, and while we
intuitively expected many children to fear that the robot would be sad if
left alone, only 2 expressed a moral position, while most of the others
actually expressed several forms of distrust towards the robot: \emph{it's
not alright to leave the robot alone, not because it is morally incorrect,
but because we do not trust the robot to remain still}. This shows that things
must be put into perspective and even young children do not consider our robot
as an intentional agent, not to say a social agent.

\begin{figure}[!h]
    \centering 
    \includegraphics[width=0.9\linewidth]{domino-leave-why.pdf}
    \caption{\small Attributions of moral
        standing: we asked children whether it was alright to leave Ranger
        alone at home if they go on a 2-week vacation. Multiple answers
        (open-ended) were allowed when asked to justify their answer.}

    \label{fig:domino-leave-alone} 
\end{figure}

\subsection{Manipulation check: Negligable impact of the conditions}
